CS6999 Web Data and Text Mining:\\Parallel Text Indexer Project Report

### 2011-06-06 report/report.tex
Finished merging results except for graphs

    an overview of the experimental design

    a detailed description of specific methods to be employed to accomplish the specific aims

    a detailed discussion of the way in which the results will be collected, analyzed, and interpreted

    a projected sequence or timetable (work plan)

    a description of any new methodology used and why it represents an improvement over the existing ones

    a discussion of potential difficulties and limitations and how these will be overcome or mitigated

    expected results, and alternative approaches that will be used if unexpected results are found

    precautions to be exercised with respect to any procedures, situations, or materials that may be hazardous to personnel or human subjects

Suggestion

@@

\bigskip
The whole idea of the project belongs in being able to reduce the time
to build the index. As our system is building the inverted index by
parsing all the post into the state machine, we can expect that the
complexity will be proportional to the whole number of characters of
all the post we want to build the index for. As a result changing the
number of Posts, the number of characters inside of each post should
have a big impact in our time complexity. Besides, we should also have
a time needed at the very beginning of the running to build the state
machine. We can expect that the building of the state machine needs a
time proportional to the sum of the size of the keywords.

@@

%----------------------------------------
% Experiment Results
%----------------------------------------
\section{Experiment Results}
\label{sec:experimentresults}
FIX: All tests were run on a Dell laptop with an Intel Core Duo 2GHz
processor, 1GB RAM, running a 32 bit Linux 2.6.37 kernel. The Java
Virtual Machine used was version 1.6.0-24.....

@@

%----------------------------------------
% Comparison with Naive Index Builder
%----------------------------------------
\subsection{Comparison with Na\"{a}ve Index Builder}
This part focuses on the improvement that our system with respect to
time complexity in comparison with a na\"{i}e index builder. The na\"{i}ve
indexer builder has the same input as our system i.e. a list of
keywords and a list of documents, and the same output i.e. an inverted
index corresponding to the keywords and documents. Algorithm
\ref{alg:invertedindex} is used in principal for both implementations.

@@

As we can see this algorithm has a complexity that directly depends on
the number of keywords n, the number of document m and the length of
document since we have to check in each document for each
keyword. Moreover, we decided to limit as much as possible the
accesses to the database since it is really time consuming in
Java. That’s why we do one access to the database for each keywords
and not for each document.

Note: In the state machine we do as much accesses to the database as
the number of document (in the worst case meaning we find at least one
keyword per document).

Let’s compare the results for this algorithm with our system results
for different number of keywords and posts.

\subsubsection{Analytical}

\subsubsection{Experimental}

%----------------------------------------
% Conclusions
%----------------------------------------
\section{Conclusions}
\label{sec:conclusions}

The experimental results tend to prove what we could expect from
theory. Our system is faster than the naive indexer whatever
parameters we are changing. In particular when the number of document
is becoming really important, the State machine approach prove to be
very efficient. Moreover, as soon as we are dealing with a number of
keywords that can be representative of industrial needs, the state
machine prove also to be far faster than the naive indexer.

Future work on this strategy would be to compare this approach with
less naive strategies to build an index. Moreover, as said in the
introduction, it would probably be interesting to add compression
methods to our algorithm in order to reduce the index size in the
database.

>>>>

\begin{itemize}
    \item an overview of the experimental design

    \item a detailed description of specific methods to be employed to accomplish the specific aims

    \item a detailed discussion of the way in which the results will be collected, analyzed, and interpreted

    \item a projected sequence or timetable (work plan)

    \item a description of any new methodology used and why it represents an improvement over the existing ones

    \item a discussion of potential difficulties and limitations and how these will be overcome or mitigated

    \item expected results, and alternative approaches that will be used if unexpected results are found

    \item precautions to be exercised with respect to any procedures, situations, or materials that may be hazardous to personnel or human subjects
\end{itemize}

Suggestion:

@@

@@

%----------------------------------------
% Experiment Results
%----------------------------------------
\section{Experiment Results}
\label{sec:experimentresults}
The whole idea of the project belongs in being able to reduce the time
to build the index. As our system is building the inverted index by
parsing all the post into the state machine, we can expect that the
complexity will be proportional to the whole number of characters of
all the post we want to build the index for. As a result changing the
number of Posts, the number of characters inside of each post should
have a big impact in our time complexity. Besides, we should also have
a time needed at the very beginning of the running to build the state
machine. We can expect that the building of the state machine needs a
time proportional to the sum of the size of the keywords.

All tests were run on a single Intel Core 2 Duo 2GHz
processor, 1GB RAM, running a 32 bit Linux 2.6.37 kernel. The Java
Virtual Machine used was version 1.6.0-24. The database used was MySQL
x.x, running on ....

@@

%----------------------------------------
% Comparison with Naive Index Builder
%----------------------------------------
\subsection{Comparison with Na\"{i}ve Index Builder}
This part focuses on the improvement that our system with respect to
time complexity in comparison with a na\"{i}e index builder. The na\"{i}ve
indexer builder has the same input as our system i.e. a list of
keywords and a list of documents, and the same output i.e. an inverted
index corresponding to the keywords and documents. Algorithm
\ref{alg:invertedindex} is used in principal for both implementations.

@@

As we can see this algorithm has a complexity that directly depends on
the number of keywords $n$, the number of document $m$, and the length
of document since we have to check in each document for each
keyword. Moreover, we decided to limit, as much as possible, the
accesses to the database since it is really time consuming in
Java. That’s why we do one access to the database for each keywords
and not for each document. In the state machine we do as much accesses
to the database as the number of document (in the worst case meaning
we find at least one keyword per document). In the following
subsections we compare the results for this algorithm with our system
results for different number of keywords and posts.

\subsubsection{Complexity Analysis}
For the na\"{i}ve algorithm described above, the theoretical worst
case time complexity relationship is shown in equation
\ref{eqn:timecomplexworstcase}.

\begin{equation}
\label{eqn:timecomplexworstcase}
Time \propto keyNum * ( docNum * maxLength + conn)
\end{equation}

where \(keyNum\) is the number of keywords, \(docNum\) is the number
of documents, \(maxLength\) is the maximum document length, and
\(conn\) is the time taken to create a connection to the database. For
the state machine algorithm the theoretical complexity relationship is
given by 
equation \ref{eqn:timecomplexsm}.

\begin{equation}
\label{eqn:timecomplexsm}
Time \propto docNum * ( maxLength + conn)
\end{equation}

\subsubsection{Document Size}
For this experiment we fixed the number of keywords to 1,000 and the
number of documents to 5,000 and we ran our code on a database
composed of document of different length. The experiment results are
shown in figure \ref{fig:naivedocumentsize}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{naivedocumentsize}
  \end{center}
  \caption{Time to build an inverted index as a function of the
      average document size. Time taken is shown in miliseconds on the
      y-axis and the average document size in characters on the x-axis}
  \label{fig:naivedocumentsize}
\end{figure}

As expected, the curves for both our indexer and the naive indexer are
almost linear since for the naive implementation:

\[Time \propto (keyNum * docNum) * maxLength + conn * keyNum\]

and for our Aho-Corasick state machine method:

\[Time \propto docNum * maxLength + conn * docNum\]

We can observe that for a small length of document naive method is
faster since \(keyNum\) is smaller than \(docNum\). For \(docLength = 1300\)
our indexer is far more efficient. To see this effect more clearly,
experiment results are plotted on a logarithmic scale in figure
\ref{fig:naivedocumentsizelog}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{naivedocumentsizelog}
  \end{center}
  \caption{Time to build an inverted index as a function of the
      average document size. Time taken is shown in miliseconds on the
      y-axis and the average document size in characters on the
      logarithmic x-axis}
  \label{fig:naivedocumentsizelog}
\end{figure}

\subsubsection{Number of Keywords}
As expected, the two implementations exhibit a linear response to the
number of keywords used. In the na\"{i}ve implmementation:

\[Time \propto (keyNum * maxLength) * docNum + conn * keyNum\]

and in our Aho-Corasick state machine method:

\[Time \propto (maxLength + conn) * docNum \]

However, we would have expected the state machine method not to depend
on the number of keyword. The reason why it’s not the case is because
when we don’t have any keywords in a document, we don’t access the
database. Knowing that the database access is very time consuming,
increasing the number of keyword increases the probability of finding a
keyword into each document and so increase the time complexity.

The experiment results are shown in figure \ref{fig:naivedocumentsize}
and on a logarithmic scale in figure \ref{fig:naivedocumentsizelog}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{naivenumkeywords}
  \end{center}
  \caption{Time to build an inverted index as a function of the
      number of keywords. Time taken is shown in miliseconds on the
      y-axis and the number of keywords on the x-axis}
  \label{fig:naivenumkeywords}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{naivenumkeywordslog}
  \end{center}
  \caption{Time to build an inverted index as a function of the
      number of keywords. Time taken is shown in miliseconds on the
      y-axis and the number of keywords on the logarithmic x-axis}
  \label{fig:naivenumkeywordslog}
\end{figure}

\subsubsection{Size of Corpus}
As expected, the time complexity is linearly related to the number of
documents being indexed. The experiment results are shown in figure
\ref{fig:naivesizecorpus} and on a logarithmic scale in figure
\ref{fig:naivesizecorpuslog}. For the na\"{i}ve implementation:

\[Time \propto (docNum * maxLenght + conn) * keyNum\]

and in our Aho-Corasick state machine method:

\[Time \propto docNum * maxLenght + conn * docNum\]

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{naivesizecorpus}
  \end{center}
  \caption{Time to build an inverted index as a function of the
      number of documents being indexed. Time taken is shown in
      miliseconds on the y-axis and the number of documents on the
      x-axis} 
  \label{fig:naivesizecorpus}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth,height=!]{naivesizecorpuslog}
  \end{center}
  \caption{Time to build an inverted index as a function of the
      number of documents being indexed. Time taken is shown in
      miliseconds on the y-axis and the number of documents on the
      logarithmic x-axis} 
  \label{fig:naivesizecorpuslog}
\end{figure}

%----------------------------------------
% Conclusions
%----------------------------------------
\section{Conclusions}
\label{sec:conclusions}
The experimental results tend to prove what we could expect from
theory. Our system is faster than the na\"{i}ve indexer with respect
to the number of keywords being indexed, size of the document
collection, and average document size. In particular, our indexer
implementation performs relatively well with respect to the number of
documents being indexed. Moreover, as soon as we are dealing with a number of
keywords that can be representative of real-world needs, the state
machine implementation proves also to be significantly faster than the
naive indexer.

One of the primary goals of this research project was to leverage the
effect of parallelization in reducing the time to index a document
collection. Due to time constraints, this was not done but the
resulting implementation lays a strong foundation for further research
in this area. Other avenues of interest include, but are not limited
to:

\begin{itemize}
  \item Compare the Aho-Corasick state machine indexer implementation
    to less na\"{i}ve indexer implementations.

    \item Use compression techniques to our algorithm in order to
      reduce the index size in the database.

    \item Study the time performance improvements possible deploying
      the Aho-Corasick state machine indexer within a commercial
      \textit{MapReduce} framework such as \textit{Hadoop}.
\end{itemize}
